Supplementary information for The transboundary nature of the world’s exploited marine species

### 2020-10-05 Scripts/Manuscript_Draft.Rmd

We have identified four key uncertainties in the method utilized that may affect the estimation of transboundary species. Firstly, as we ran the analysis at the species level due to limited spatial-specific data on species sub populations (stocks), we are not able to identify transboundary stocks within EEZs. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research suggests strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca]. **However, it is important to acknowledge that there is considerable uncertainty associated in determining levels of connectivity across time and space for marine populations at different life stages, from larvae [@Kaplan:2016hs] to adults [@Archambault:2016ie]. This is of special concern, but-not-limited-to [@Archambault:2016ie], highly migratory species whose ranges span multiple jurisdictions and the high seas, such as tunas [@Moore:2020cd], billfishes [@Sepulveda:2019hq] and sharks [@Vandeperre:2014jk] challenging management decisions based on meta-populations [@Moore:2020cd].** Here, we limited the definition of “transboundary” to include species spanning only adjacent countries (e.g., US and Canada), excluding countries that were separated by another nation (e.g., Canada and Mexico) and/or the high seas (e.g., Canada and Spain). Consequently, our results likely provide a conservative estimate of transboundary species, as we did not cover all marine taxa in the world [@Reygondeau:2019uh] and only analyzed species present in all four data sources. Nevertheless, our results are representative of a substantial proportion of the world's marine catches and revenue from economically important marine fisheries. Secondly, the predicted species' distribution is affected by the uncertainties of the environmental data and models used for such predictions. Structural differences within Earth System Models result in variations of oceanic conditions, which undoubtedly affect the ENM-Nereus. We averaged results from the three models in an effort to capture the structural variation across models. Natural climate variability is a major driver of marine species' distributions, potentially removing a species from an EEZ for long periods (e.g. Anchovies and Sardines "regimes" in the Eastern Pacific are strongly influenced by water temperature decal oscillations [@Chavez:2003jr]). Thus, a species’ distribution can potentially be reduced or shifted in such a way that it only covers one EEZ until oceanic conditions change again and the associated species' distribution expands. To account for such climate variability, we derived the ENM-Nereus results as an average of oceanic conditions between 1970 and 2000. This is not an issue for the SDM-SAU as it does not directly require environmental variables [@Pauly:2016uz]. Thirdly, we assumed that, if the*Sea Around Us* reconstructed data recorded a species as caught in any given grid cell, then the species was present within that grid cell. While catch data are not exempt of uncertainty, in most cases, differences between the*Sea Around Us* and the FAO self-reported data are smaller towards the end of the time series [@Pauly:2019bf]. Thus, we limited the catch data reference period in our analyses to between 2005 and 2014, the last ten years of data available. Likewise, the spatial allocation of the catch is subjected to imprecision, mainly produced by differences in the spatial scale of the original data and the method employed by the*Sea Around Us*. Finally, this study’s results only considered species for which all datasets agreed on presence and had and *Area Index* of at least 25% (e.g., the species shared distribution was at least 25% in each EEZ). Therefore, again, our approach presents a relatively conservative estimate of the number of transboundary species in the world. Using a smaller *Area Index* will result in more transboundary species (Supplementary Fig. S2).

>>>>

We have identified four key uncertainties in the method utilized that may affect the estimation of transboundary species. Firstly, as we ran the analysis at the species level due to limited spatial-specific data on species sub populations (stocks), we are not able to identify transboundary stocks within EEZs. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research suggests strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca]. **However, it is important to acknowledge that there is considerable uncertainty associated in determining levels of connectivity across time and space for marine populations at different life stages, from larvae [@Kaplan:2016hs] to adults [@Archambault:2016ie]. This is of special concern, but-not-limited-to [@Archambault:2016ie], highly migratory species whose ranges span multiple jurisdictions and the high seas, such as tunas [@Moore:2020cd], billfishes [@Sepulveda:2019hq] and sharks [@Vandeperre:2014jk] challenging management decisions based on meta-populations [@Moore:2020cd].** Here, we limited the definition of “transboundary” to include species spanning only adjacent countries (e.g., US and Canada), excluding countries that were separated by another nation (e.g., Canada and Mexico) and/or the high seas (e.g., Canada and Spain). Consequently, our results likely provide a conservative estimate of transboundary species, as we did not cover all marine taxa in the world [@Reygondeau:2019uh] and only analyzed species present in all four data sources. Nevertheless, our results are representative of a substantial proportion of the world's marine catches and revenue from economically important marine fisheries. Secondly, the predicted species' distribution is affected by the uncertainties of the environmental data and models used for such predictions. Structural differences within Earth System Models result in variations of oceanic conditions, which undoubtedly affect the ENM-Nereus. We averaged results from the three models in an effort to capture the structural variation across models. Natural climate variability is a major driver of marine species' distributions, potentially removing a species from an EEZ for long periods (e.g. Anchovies and Sardines "regimes" in the Eastern Pacific are strongly influenced by water temperature decal oscillations [@Chavez:2003jr]). Thus, a species’ distribution can potentially be reduced or shifted in such a way that it only covers one EEZ until oceanic conditions change again and the associated species' distribution expands. To account for such climate variability, we derived the ENM-Nereus results as an average of oceanic conditions between 1970 and 2000. This is not an issue for the SDM-SAU as it does not directly require environmental variables [@Pauly:2016uz]. Thirdly, we assumed that, if the *Sea Around Us* reconstructed data recorded a species as caught in any given grid cell, then the species was present within that grid cell. While catch data are not exempt of uncertainty, in most cases, differences between the*Sea Around Us* and the FAO self-reported data are smaller towards the end of the time series [@Pauly:2019bf]. Thus, we limited the catch data reference period in our analyses to between 2005 and 2014, the last ten years of data available. Likewise, the spatial allocation of the catch is subjected to imprecision, mainly produced by differences in the spatial scale of the original data and the method employed by the *Sea Around Us*. Finally, this study’s results only considered species for which all datasets agreed on presence and had and *Area Index* of at least 25% (e.g., the species shared distribution was at least 25% in each EEZ). Therefore, again, our approach presents a relatively conservative estimate of the number of transboundary species in the world. Using a smaller *Area Index* will result in more transboundary species (Supplementary Fig. S2).

### 2020-10-05 Scripts/supplements.Rmd

---
title: "Supplementary information for Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes, Gabriel Reygondeau, Colette Wabnitz and William W.L. Cheung"
output: pdf_document
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
- \floatplacement{table}{H}
---

>>>>

---
title: "Supplementary information for The transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes, Gabriel Reygondeau, Colette C.C. Wabnitz and William W.L. Cheung"
output: pdf_document
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
- \floatplacement{table}{H}
---
